### 2025-02-24 paper/relatedwork.tex
More.

\subsection{Type systems}
Verifying index operations is a classic application of dependent
types, as seen in for example Dependent
ML~\cite{10.1145/292540.292560}, although most work assumes static
ranks. Our target language, Futhark, supports size-dependent
types~\cite{10.1145/3609024.3609412}, but this information is used
only to ensure shape conformance (e.g., that the operands to a matrix
multiplication have the same size), not to ensure the correctness of
array indexing.

Gibbons showed how to express rank polymorphism in Haskell, through
the use of \emph{Naperian functors}~\cite{10.1145/2976022.2976023}.
This form of rank polymorphism is somewhat more limited than ours; in
particular, it imposes static ranks. Further, Gibbons' work focused
solely on the embedding in Haskell, while our system also supports
code extraction and demonstrates good real-world performance.

>>>>

\subsection{Type systems}

Verifying index operations is a classic application of dependent
types, as seen in for example Dependent
ML~\cite{10.1145/292540.292560}, although most work assumes static
ranks. Our target language, Futhark, supports size-dependent
types~\cite{10.1145/3609024.3609412}, but this information is used
only to ensure shape conformance (e.g., that the operands to a matrix
multiplication have the same size), not to ensure the correctness of
array indexing. Single-Assignment C~\cite{sac2} (SaC) is a statically
typed language with support for rank polymorphism, including dynamic
ranks, but it does not guarantee that absence of indexing errors.

Gibbons showed how to express rank polymorphism in Haskell, through
the use of \emph{Naperian functors}~\cite{10.1145/2976022.2976023}.
This form of rank polymorphism is somewhat more limited than ours; in
particular, it imposes static ranks. Further, Gibbons' work focused
solely on the embedding in Haskell, while our system also supports
code extraction and demonstrates good real-world performance.
Similarly, Remora~\cite{rank-poly} is a dependently typed
rank-polymorphic language that also supports only static ranks.
